\chapter*{Abstract \label{sec:abstract}}
\addcontentsline{toc}{chapter}{Abstract}

Ensuring compiler correctness is critical to mitigating the risks of errors, especially during the optimization phases of a compiler. 
Despite existing approaches to identifying incorrect optimizations in GraalVM's expression optimizations, none incorporate a solver-based tool 
that provides GraalVM developers with a sufficiently comprehensive analysis of the optimization rules without requiring expertise in formal proofs. 
Thus, we introduce VeriTest, a tool designed to automatically verify the existence of a simple proof of optimization correctness or identify 
counterexamples for incorrect optimizations. VeriTest leverages Isabelle's automated tools and Veriopt's representation of GraalVM's intermediate 
representation (IR) as a domain-specific language (DSL) within Isabelle. This combination enables a comprehensive analysis of optimization rules. 
The implementation abstracts Isabelle Client - Server interactions to ensure efficient and concurrent processing of optimization rules, while also 
assuring extensibility for future refinements. Our evaluation demonstrates the capabilities of VeriTest in automatically verifying optimization 
rules, though significant limitations were found in identifying counterexamples due to the inherent limitations of Isabelle and Veriopt's DSL. However, 
this also implies that the analysis is expected to improve as the theoretical foundations for GraalVM's IR is improved.
Despite these challenges, VeriTest has shown significant potential as an automatic verification tool for GraalVM's expression optimizations, 
providing a solid foundation for future enhancements.